User-customized hashtag/tag prediction on Twitter and StackOverflow:\protect \\ Comparing vector-based and ACT-R Bayesian models of tag retrieval

### 2013-10-06 writeups/proposal/proposal.tex
foo
Former-commit-id: 5d39450145900210f2d0354e3ade31cca54726a0 [formerly 3b51d3b610e8e700a938b7552e75905b1a9775a5]
Former-commit-id: 1fec6f77c529e831d6c424e45b944a7134a80c15

\subsubsection{StackOverflow and Twitter Task Domains}

@@

The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across different task domains.
However, they are also similar on several accounts:
Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags and tags for Twitter and StackOverflow respectively),
and the user data from the datasets are publicly available for analysis.

Also, studying user-based tag generation on these sites is a relevant task domain.
Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
These systems can also help newer users by recommending proper tags for their newest content.
On the StackOverflow site for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
For Twitter, hashtags represent streams of information that are possibly interesting to the user.
Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams.

@@

\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
However, this model scaled much better than the co-occurrence model presented in \textcite{Kuo2011}.
Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
After scaling up the model, this model can successfully predict a user's tag for a post at 65\% accuracy.
It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is built.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular task domain:
predicting the user-chosen tags for human-created content on social media sites.

@@

The end result of this research will be a better understanding of how a user's past tagging history influences future tag use,
as well as two cognitively-plausible models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared on a real-world tagging task, which has never been done previously.
The strengths of each model (word order and user prior likelihood) will be incorporated into the other, and performance differences will be evaluated.
The main two proposed model modifications (incorporating word order into ACT-R and user history into vector-based representations) have not been tested before, and seem likely to significantly improve performance.
Further, these two modifications in particular are aimed at the architectural level of each model, may very well be task independent, and may generalize and improve model fit across other tasks.
Finally, if model accuracy is high enough, then the developed models may be applied to general tag recommendation systems for either Twitter or StackOverflow,
and possibly for other sites where users generate and tag content.

>>>>

\subsubsection{StackOverflow and Twitter}

@@

The StackOverflow and Twitter datasets were chosen because the human-created content between them is quite different, and I am interested in retrieval models that generalize across tasks.
However, they are also similar on several accounts:
Both domains have amassed large amounts of user-created content, users generate tags for posts when creating content on the site (hashtags and tags for Twitter and StackOverflow respectively),
and the user data from the datasets are publicly available for analysis.

Also, studying user-based tag generation on these sites has relevant real-world application.
Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
These systems can also help newer users by recommending proper tags for their newest content.
On the StackOverflow site for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
For Twitter, hashtags represent streams of information that are possibly interesting to the user.
Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams.

@@

\textcite{Stanley2013} also modeled a user's chosen tags for a post on the StackOverflow site.
The model was based on ACT-R's declarative memory retrieval theory, which at the core is also a Bayesian co-occurrence model.
However, this model scaled much better than the co-occurrence model presented in \textcite{Kuo2011}.
Because of this, 100 times more posts were used to build the co-occurrence matrix (\num{1000000} versus \num{10000}), which significantly increased the model's performance.
After scaling up the model, this model can successfully predict a user's tag for a post at 65\% accuracy.
It is interesting that in both research papers the best performing model was based on a Bayesian co-occurrence framework, since that is precisely the framework from which ACT-R's declarative memory system is based.
This suggests that more cognitively-plausible memory retrieval models may outperform more general machine-learning approaches in this particular environment
(i.e., when predicting the user-chosen tags for human-created content on social media sites).

@@

The end result of this research will be a better understanding of how a user's past tagging history influences future tag use,
as well as two cognitively-plausible models for making user-customized tag predictions for StackOverflow posts and Twitter tweets. 
Overall performance differences between an ACT-R's Bayesian retrieval model and a random-permutation vector-based model will be directly compared on two real-world tagging tasks, which has never been done previously.
The strengths of each model (word order and user prior likelihood) will be incorporated into the other, and performance differences will be evaluated.
The main two proposed model modifications (incorporating word order into ACT-R and user history into vector-based representations) have not been tested before, and seem likely to significantly improve performance.
Further, these two modifications in particular are aimed at the architectural level of each model, may very well be task independent, and may generalize and improve model fit across other tasks.
Finally, if model accuracy is high enough, then the developed models may be applied to general tag recommendation systems for either Twitter or StackOverflow,
and possibly for other sites where users generate and tag content.
